Figures for Paper

### 2021-03-19 paper_figures.Rmd
Update paper_figures.Rmd
fix git merge flags

## barplot UNC13B NMD PSI

@@

## barplot UNC13A normalized IR
```{r}
unc13a_ir = rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "normalized_unc13a_ir",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("Normalized UNC13A \nIntron Retention Ratio") + 
    xlab("") + 
    theme(legend.position = "none") +
    guides(color = FALSE) + 
      theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))
```

@@

rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "TARDBP",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("TARDBP") + 
    xlab("") + 
    theme(legend.position = "none") +
    guides(color = FALSE) + 
      theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28)) +
    scale_y_continuous(labels = scales::percent) +
    geom_text(data = text_table[gene_name == "TARDBP"],aes(x = source, y= 1.35,label = padj_plot),size = 8) + 
    geom_text(data = text_table[gene_name == "TARDBP"],aes(x = source, y= 1.2,label = log2FoldChange_plot),size = 8)

@@

## Scatterplot UNC13B-IR and UNC13A Crptic

```{r}
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = cryptic_psi_full, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = cryptic_psi_full, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')+ 
  facet_wrap(~source)
```

@@

nmd_plot = ggbarplot(for_bar_plot_sy,
                        x = 'gene',
                        add = c("mean_se","jitter"),
                        y = 'value',
                        fill = 'condition',
                        color = 'condition',
                        position = position_dodge(0.8),
                        dot.size = 10) + 
  scale_y_continuous(labels = scales::percent) + 
  ggpubr::theme_pubr() +
  scale_fill_manual(
    values = c("#40B0A6","#E1BE6A")
  ) + 
  scale_color_manual(
    values = c("#1C2617","#262114")
  ) +
  ylab("Percent of transcript after CHX treatment")

@@

```{r}
clean_data_table = fread(file.path(here::here(),"data","nygc_junction_information.csv"))
clean_data_table = clean_data_table %>%     
    mutate(rs12973192 = fct_relevel(rs12973192,
                              "C/C", "C/G", "G/G")) %>% 
    mutate(number_g_alleles = as.numeric(rs12973192) - 1) %>% 
    mutate(unc13a_cryptic_leaf_psi = ifelse(is.na(unc13a_cryptic_leaf_psi),0,unc13a_cryptic_leaf_psi))

@@

```{r}
comps = unique(clean_data_table$disease_group2) %>% gsub("Control", " Control",.) %>% 
  combn(.,2,simplify = F)

clean_data_table %>% 
  group_by(disease_group2,tissue_clean) %>% 
  mutate(n_sample = n_distinct(sample)) %>% 
  ungroup() %>% 
  mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>% 
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
 filter(tissue_clean %in% 
           c("Frontal Cortex", 
             "Lumbar Spinal Cord", 
             "Cervical Spinal Cord",
             "Motor Cortex", 
             "Temporal Cortex",
             "Cerebellum") )%>% 
mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                              "Frontal Cortex",
                              "Lumbar Spinal Cord",
                              "Temporal Cortex",
                              "Motor Cortex")) %>% 
    ggplot(aes(x = detection_name, y = UNC13A_TPM)) +
    ggpubr::theme_pubr() +
    geom_boxplot(aes(fill = disease_group2)) + 
    geom_jitter(pch = 21, color = 'black',width  = 0.2,height = 0,aes(fill =   disease_group2)) +
    ylab("UNC13A TPM") + 
    xlab("") + 
    guides(color = FALSE) +
    theme(text = element_text(size = 20)) +
  facet_wrap(~tissue_clean, scales = 'free',nrow = 3) +
  theme(legend.position = 'none')
```

@@

```{r}
clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(unc13a_cryptic_leaf_psi > 0) %>% 
  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = UNC13B_TPM)) + 
  geom_point() + 
  stat_cor(label.y = 38,size = 8,method = 's',cor.coef.name = 'rho') + 
  geom_smooth(method = lm, se = FALSE,color = "black") +
  ylab("UNC13A TPM") +
  xlab("UNC13A CE PSI") + 
  ggpubr::theme_pubr() + 
  theme(text = element_text(size = 22)) +
  scale_x_continuous(labels = scales::percent)

@@

```{r}
clean_data_table %>% 
  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
    filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = UNC13A_TPM, color = disease_group2)) + 
  geom_point() + 
  stat_cor(size = 8) + 
  geom_smooth(method = lm, se = FALSE,color = "black") +
  ylab("UNC13A TPM") +
  xlab("UNC13A CE PSI") + 
  ggpubr::theme_pubr() + 
  theme(text = element_text(size = 18)) +
  facet_wrap(~tissue_clean,scales = 'free') +
  scale_x_continuous(labels = scales::percent)

@@

```{r}
clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
  filter(unc13a_cryptic_leaf_psi > 0) %>% 
  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = unc13a_cryptic_leaf_psi)) + 
  geom_point() + 
  stat_cor(size = 8,method = "spearman",cor.coef.name = "rho") + 
  geom_smooth(method = lm, se = FALSE,color = "black") +
  xlab("STMN2 Cryptic PSI") +
  ylab("UNC13A CE PSI") + 
  ggpubr::theme_pubr() + 
  theme(text = element_text(size = 18)) +
  scale_y_continuous(labels = scales::percent)  +
  scale_x_continuous(labels = scales::percent) 
```

## Side by side comparison UNC13A & STMN2 CE PSI
```{r}
clean_data_table %>% 
  # filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
  # filter(unc13a_cryptic_leaf_psi > 0) %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  melt() %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
    mutate(tissue_clean = fct_relevel(tissue_clean,"Motor Cortex", "Cervical Spinal Cord",
                                    "Frontal Cortex",
                                    "Lumbar Spinal Cord",
                                    "Temporal Cortex")) %>% 
  mutate(variable = ifelse(grepl("stmn_2",variable), "STMN2", "UNC13A")) %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot(aes(color = variable)) +
  geom_jitter(aes(fill = variable),pch = 21,height = 0,width = 0.2,size = 3) +
  scale_fill_manual(values = c("#008574","#cd5582")) +
  scale_color_manual(values = c("#008574","#cd5582")) +
  facet_wrap(~tissue_clean,nrow = 3,scales = 'free') +
    scale_y_continuous(labels = scales::percent)  +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24))  +
  xlab("") +
  ylab("PSI")  +
  theme(legend.position = 'none')

@@

# Effect of UNC13A genotype
## UNC13A psi by genotype - rs12973192
```{r}
rs12973192_comparisons = clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  group_by(rs12973192) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
  pull(detection_name) %>% 
  unique() %>% 
  combn(.,2,simplify = F)

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  group_by(rs12973192) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = detection_name, fill = rs12973192)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  ylab("UNC13A PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means() +
  stat_compare_means(comparisons = rs12973192_comparisons,label = "p.signif") + 
  scale_y_continuous(trans='log10')
```

## UNC13A psi by genotype - rs12608932
```{r}

rs12608932_comparisons = clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  group_by(rs12608932) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{rs12608932} \n ( {n_sample} )")) %>% 
  pull(detection_name) %>% 
  unique() %>% 
  combn(.,2,simplify = F)

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  group_by(rs12608932) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{rs12608932} \n ( {n_sample} )")) %>% 
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = detection_name, fill = rs12608932)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  ylab("UNC13A PSI") +
  xlab("UNC13A rs12608932 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means() +
  stat_compare_means(comparisons = rs12608932_comparisons,label = "p.signif") + 
  scale_y_continuous(trans='log10')

@@

@@

## Correlation UNC13A PSI and STMN2 PSI with genotype effect

@@

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
    mutate(rs12973192 = fct_relevel(rs12973192,
                              "C/C", "C/G", "G/G")) %>%
  filter(!grepl("Cord",tissue_clean)) %>% 
  mutate(log10_stmn2 = log10(stmn_2_cryptic_psi_leaf)) %>% 
    mutate(log10_unc13a = log10(unc13a_cryptic_leaf_psi)) %>% 
    ggpubr::ggscatter(., 
                      x = "log10_stmn2",
                      y = "log10_unc13a",
                      color = 'rs12973192',                      
                      add = "reg.line", size = 3
                      ) +
    ylab("UNC13A CE PSI ") + 
    xlab("STMN2 Cryptic PSI ") +
  stat_cor(aes( x = stmn_2_cryptic_psi_leaf,
                      y = unc13a_cryptic_leaf_psi,
                      color = rs12973192),
           method = 'spearman',
           cor.coef.name = 'rho',
           show.legend = FALSE,
           size = 14) + 
    scale_color_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  theme(legend.title = element_blank()) +
  theme(text = element_text(size = 24),legend.text = element_text(size = 32))

@@

@@

>>>>

## barplot UNC13B frameshift PSI

@@

## barplot UNC13A normalized IR
```{r}
unc13a_ir = rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "normalized_unc13a_ir",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("Normalized UNC13A \nIntron Retention Ratio") + 
    xlab("") + 
    theme(legend.position = "none") +
    guides(color = FALSE) + 
      theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))

@@

rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "TARDBP",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("TARDBP") + 
    xlab("") + 
    theme(legend.position = "none") +
    guides(color = FALSE) + 
      theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28)) +
    scale_y_continuous(labels = scales::percent,expand = expansion(mult = c(0, .1))) +
    geom_text(data = text_table[gene_name == "TARDBP"],aes(x = source, y= 1.35,label = padj_plot),size = 8) + 
    geom_text(data = text_table[gene_name == "TARDBP"],aes(x = source, y= 1.2,label = log2FoldChange_plot),size = 8)

@@

## Scatterplot UNC13B-IR and UNC13A Cryptic

rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = cryptic_psi_full, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = cryptic_psi_full, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')+ 
  facet_wrap(~source)
```

@@

nmd_plot = ggbarplot(for_bar_plot_sy,
                        x = 'condition',
                        add = c("mean_se","jitter"),
                        y = 'value',
                        fill = 'condition',
                        color = 'condition',
                        position = position_dodge(0.8),
                        dot.size = 10,facet.by = 'gene') + 
  scale_y_continuous(labels = scales::percent) + 
  ggpubr::theme_pubr() +
  scale_fill_manual(
    values = c("#40B0A6","#E1BE6A")
  ) + 
  scale_color_manual(
    values = c("#1C2617","#262114")
  ) +
  ylab("Percent of transcript after CHX treatment") + 
  stat_compare_means()

@@

```{r}
clean_data_table = fread(file.path(here::here(),"data","nygc_junction_information.csv"))
clean_data_table = clean_data_table %>%     
    mutate(rs12973192 = fct_relevel(rs12973192,
                              "C/C", "C/G", "G/G")) %>% 
    mutate(number_g_alleles = as.numeric(rs12973192) - 1) %>% 
    mutate(unc13a_cryptic_leaf_psi = ifelse(is.na(unc13a_cryptic_leaf_psi),0,unc13a_cryptic_leaf_psi)) %>% 
    mutate(junction_reads_stmn2 = STMN2_annotated_leaf + STMN2_cryptic_leaf) %>% 
  mutate(junction_reads_unc13a = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf + UNC13A_annotated_leaf) %>% 
  unique()

@@

```{r}
comps = clean_data_table %>% 
  group_by(disease_group2,tissue_clean) %>% 
  mutate(n_sample = n_distinct(sample)) %>% 
  ungroup() %>% 
  mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>% 
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% pull(detection_name) %>% 
  unique() %>% 
  combn(.,2,simplify = F)

clean_data_table %>% 
  group_by(disease_group2,tissue_clean) %>% 
  mutate(n_sample = n_distinct(sample)) %>% 
  ungroup() %>% 
  mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>% 
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
 filter(tissue_clean %in% 
           c("Frontal Cortex", 
             "Lumbar Spinal Cord", 
             "Cervical Spinal Cord",
             "Motor Cortex", 
             "Temporal Cortex",
             "Cerebellum") )%>% 
mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                              "Frontal Cortex",
                              "Lumbar Spinal Cord",
                              "Temporal Cortex",
                              "Motor Cortex")) %>% 
    ggplot(aes(x = detection_name, y = UNC13A_TPM)) +
    ggpubr::theme_pubr() +
    geom_boxplot(aes(fill = disease_group2)) + 
    geom_jitter(pch = 21, color = 'black',width  = 0.2,height = 0,aes(fill =   disease_group2)) +
    ylab("UNC13A TPM") + 
    xlab("") + 
    guides(color = FALSE) +
    theme(text = element_text(size = 20)) +
  facet_wrap(~tissue_clean, scales = 'free',nrow = 3) +
  theme(legend.position = 'none') + 
  stat_compare_means(comparisons = comps)
```

```{r make_those_plots_function}
plot_by_tissue = function(tissue_name,dt){

only_tissue = dt %>% 
  group_by(disease_group2,tissue_clean) %>% 
  mutate(n_sample = n_distinct(sample)) %>% 
  ungroup() %>% 
  filter(tissue_clean == tissue_name) %>% 
  mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>% 
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean))

unique_x_values = only_tissue %>% 
  pull(detection_name) %>% 
  unique()

comparisons = combn(unique_x_values,2,simplify = F)

the_plot = only_tissue %>% 
    ggplot(aes(x = detection_name, y = UNC13A_TPM)) +
    ggpubr::theme_pubr() +
    geom_boxplot(aes(fill = disease_group2)) + 
    geom_jitter(pch = 21, color = 'black',width  = 0.2,height = 0,aes(fill =   disease_group2)) +
    ylab("UNC13A TPM") + 
    xlab("") + 
    guides(color = FALSE) +
    theme(text = element_text(size = 20)) +
  facet_wrap(~tissue_clean, scales = 'free',nrow = 3) +
  theme(legend.position = 'none') + 
  stat_compare_means() + 
  stat_compare_means(label = "p.signif",comparisons = comparisons,)

  return(the_plot)
}

dis_tis = clean_data_table[disease_tissue == T,unique(tissue_clean)]
dis_tis = c("Cerebellum",dis_tis)
my_plots = purrr::map(dis_tis,plot_by_tissue, clean_data_table)
ggarrange(plotlist=my_plots)

@@

```{r}
clean_data_table %>% 
  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
  filter(UNC13A_annotated_leaf > 15) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(unc13a_cryptic_leaf_psi > 0) %>% 
  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = UNC13B_TPM)) + 
  geom_point() + 
  stat_cor(label.y = 20,size = 8,method = 's',cor.coef.name = 'rho') + 
  geom_smooth(method = lm, se = FALSE,color = "black") +
  ylab("UNC13A TPM") +
  xlab("UNC13A CE PSI") + 
  ggpubr::theme_pubr() + 
  theme(text = element_text(size = 22)) +
  scale_x_continuous(labels = scales::percent)

@@

```{r}
clean_data_table %>% 
  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
  filter(UNC13A_annotated_leaf > 15) %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = UNC13A_TPM, color = disease_group2)) + 
  geom_point() + 
  stat_cor(size = 8) + 
  geom_smooth(method = lm, se = FALSE,color = "black") +
  ylab("UNC13A TPM") +
  xlab("UNC13A CE PSI") + 
  ggpubr::theme_pubr() + 
  theme(text = element_text(size = 18)) +
  facet_wrap(~tissue_clean,scales = 'free') +
  scale_x_continuous(labels = scales::percent)

@@

```{r}
clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  filter(junction_reads_stmn2 >= 30) %>% 
  filter(junction_reads_unc13a >= 30) %>%
  # filter(UNC13A_3prime_leaf + UNC13A_5prime_1_leaf >= 2) %>% 
  # filter(STMN2_cryptic_leaf >= 2) %>% 
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
  filter(unc13a_cryptic_leaf_psi > 0) %>% 
  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = unc13a_cryptic_leaf_psi)) + 
  geom_point() + 
  stat_cor(size = 8,method = "spearman",cor.coef.name = "rho") + 
  geom_smooth(method = lm, se = FALSE,color = "black") +
  xlab("STMN2 Cryptic PSI") +
  ylab("UNC13A CE PSI") + 
  ggpubr::theme_pubr() + 
  theme(text = element_text(size = 18)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1L))  
```

## Side by side comparison UNC13A & STMN2 CE PSI
```{r}
clean_data_table %>% 
    mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
    filter(tissue_clean %in% 
               c("Frontal Cortex", 
                 "Lumbar Spinal Cord", 
                 "Cervical Spinal Cord",
                 "Motor Cortex", 
                 "Temporal Cortex",
                 "Cerebellum") )%>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  filter(disease_tissue == T) %>% 
  filter(UNC13A_3prime_leaf + UNC13A_5prime_1_leaf >= 1) %>% 
  filter(STMN2_cryptic_leaf >= 1) %>% 
  dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  melt() %>% 
    mutate(tissue_clean = fct_relevel(tissue_clean,"Motor Cortex", "Cervical Spinal Cord",
                                    "Frontal Cortex",
                                    "Lumbar Spinal Cord",
                                    "Temporal Cortex")) %>% 
  mutate(variable = ifelse(grepl("stmn_2",variable), "STMN2", "UNC13A")) %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot(aes(color = variable),outlier.shape = NA) +
  geom_jitter(aes(fill = variable),pch = 21,height = 0,width = 0.2,size = 3) +
  scale_fill_manual(values = c("#008574","#cd5582")) +
  scale_color_manual(values = c("#008574","#cd5582")) +
  facet_wrap(~tissue_clean,nrow = 3, scales = 'free') +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24))  +
  xlab("") +
  ylab("CE PSI")  +
  theme(legend.position = 'none')

@@

# Effect of UNC13A genotype

## UNC13A psi by genotype - rs12973192 - RE

```{r}
rs12973192_comparisons = clean_data_table %>% 
  filter(junction_reads_unc13a >=30) %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  group_by(rs12973192) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )"))

comp_for_gg = rs12973192_comparisons %>% pull(detection_name) %>% 
  unique() %>% 
  combn(.,2,simplify = F)

ce_psi = rs12973192_comparisons %>% 
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = detection_name, fill = rs12973192)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  ylab("UNC13A PSI") +
  xlab(element_blank()) +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means() +
  theme(legend.position = "none") +
  stat_compare_means(comparisons = comp_for_gg,label = "p.signif") + 
  scale_y_continuous(trans='log10',labels = scales::percent_format(accuracy = 1L))  +
  ggtitle("CE SNP (rs12973192)")

@@

## UNC13A psi by genotype - rs12608932 - RI
```{r}
rs12608932_comparisons = clean_data_table %>% 
  filter(junction_reads_unc13a >=30) %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  group_by(rs12608932) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{rs12608932} \n ( {n_sample} )"))

comp_for_gg = rs12608932_comparisons %>% pull(detection_name) %>% 
  unique() %>% 
  combn(.,2,simplify = F)

in_psi = rs12608932_comparisons %>% 
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = detection_name, fill = rs12608932)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  ylab(element_blank()) +
  xlab(element_blank()) +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(legend.position = "none") +
  theme(text = element_text(size = 24)) +
  stat_compare_means() +
  stat_compare_means(comparisons = comp_for_gg,label = "p.signif") + 
    scale_y_continuous(trans='log10',labels = scales::percent_format(accuracy = 1L))  +
  ggtitle("intronic SNP (rs12973192)")
```

```{r}
ggpubr::ggarrange(ce_psi,in_psi)
```

@@

## Correlation UNC13A PSI and STMN2 PSI with genotype effect - rs12973192

@@

ce_snp = clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(junction_reads_stmn2 >= 30) %>% 
  filter(junction_reads_unc13a >= 30) %>%
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
    mutate(rs12973192 = fct_relevel(rs12973192,
                              "C/C", "C/G", "G/G")) %>%
  filter(!grepl("Cord",tissue_clean)) %>% 
  mutate(log10_stmn2 = log10(stmn_2_cryptic_psi_leaf)) %>% 
    mutate(log10_unc13a = log10(unc13a_cryptic_leaf_psi)) %>% 
    ggpubr::ggscatter(., 
                      x = "log10_stmn2",
                      y = "log10_unc13a",
                      color = 'rs12973192',                      
                      add = "reg.line", size = 3
                      ) +
    ylab("UNC13A CE PSI ") + 
    xlab("STMN2 Cryptic PSI ") +
  stat_cor(aes( x = stmn_2_cryptic_psi_leaf,
                      y = unc13a_cryptic_leaf_psi,
                      color = rs12973192),
           method = 'spearman',
           cor.coef.name = 'rho',
           show.legend = FALSE,
           size = 14) + 
    scale_color_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  theme(legend.title = element_blank()) +
  theme(text = element_text(size = 24),legend.text = element_text(size = 32)) +
    ylab(expression(paste("Lo", g[10], "UNC13A CE PSI"))) +
      xlab(expression(paste("Lo", g[10], "STMN2 CE PSI"))) +
  ggtitle("CE SNP (rs12973192) genotype")

@@

## Correlation UNC13A PSI and STMN2 PSI with genotype effect - rs12608932

in_snp = clean_data_table %>% 
  filter(disease_tissue == T) %>% 
    filter(junction_reads_stmn2 >= 30) %>% 
  filter(junction_reads_unc13a >= 30) %>%
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
    mutate(rs12608932 = fct_relevel(rs12608932,
                              "A/A", "A/C", "C/C")) %>%
  filter(!grepl("Cord",tissue_clean)) %>% 
  mutate(log10_stmn2 = log10(stmn_2_cryptic_psi_leaf)) %>% 
    mutate(log10_unc13a = log10(unc13a_cryptic_leaf_psi)) %>% 
    ggpubr::ggscatter(., 
                      x = "log10_stmn2",
                      y = "log10_unc13a",
                      color = 'rs12608932',                      
                      add = "reg.line", size = 3
                      ) +
    ylab("UNC13A CE PSI ") + 
    xlab("STMN2 Cryptic PSI ") +
  stat_cor(aes( x = stmn_2_cryptic_psi_leaf,
                      y = unc13a_cryptic_leaf_psi,
                      color = rs12608932),
           method = 'spearman',
           cor.coef.name = 'rho',
           show.legend = FALSE,
           size = 14) + 
    scale_color_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  theme(legend.title = element_blank()) +
  theme(text = element_text(size = 24),legend.text = element_text(size = 32)) +
      ylab(expression(paste("Lo", g[10], "UNC13A CE PSI"))) +
      xlab(expression(paste("Lo", g[10], "STMN2 CE PSI"))) +
   ggtitle("intronic SNP (rs12608932) genotype")

```{r}
ggpubr::ggarrange(ce_snp,in_snp)
```

@@

detected_table_for_graph %>% 
  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) + 
  geom_boxplot(show.legend = F) + 
  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) + 
  stat_compare_means() + 
  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
  ylab("Ratio UNC13A CE PSI / STMN2 PSI") + 
  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  geom_hline(yintercept = 0,size = 1.5) 
```
